\section{Introduction}



\section{Static and exogenous models}

The main question of the book from \cite{sutton_91} revolves around how do markets evolve to be less or more concentrated? Moreover, Sutton tries to explain why advertising is higher in concentrated industries. In order to explore the answers to these questions, he looks at different market structures and analyzes what makes these structure coherent with the data.

\subsection{Perfect competition}

Consider a market with perfect competition (price-taking assumption), exogenous sunk costs and free entry. The minimum efficient level of production (in the long run) is where $p = \min AC $. 

This assumption means that, in the long run, firms will produce only the quantity satisfying this assumption, not more, not less. This implies that the number of firms in a market only depends on the size of the market, $M$. In particular, if you denote the quantity produced by firms at the optimum as $q^*$, you will have that the optimal number of firms $n^*$ is given simply by: $$n^* = M/q^* $$
As $M\to\infty$, we also have $n^*\to\infty$, thus the concentration ratio will tend to 0.

The issue with this simplistic model is that it might not hold in many settings where (1) competition might be imperfect or (2) sunk costs might be endogenous. As an example consider the increase of population in the United States since the creation of Pepsi and Coca-Cola. Even though the population has more than doubled, their respective market shares are still about 30\% and 60\%, which clearly disproves the previous model. For that reason he explores both directions.

\subsection{Imperfect competition}

Now suppose you allow for imperfect competition. For that purpose, we consider a two-stage game where in the first stage, firms choose to enter a market and incur a cost of $A$; then in the second stage, firms play the market game.

Moreover, assume that all $M$ consumers have an income of 1 to spend exclusively on the good produced by the firms, such that in equilibrium: $$p\cdot Q = M$$
Finally, suppose the marginal cost of production is $c>0$.

\subsubsection{Bertrand competition}



\subsubsection{Cournot competition}

In the Cournot solution, assuming $N$ firms enter, a firm $i$ will choose its quantity $q_i$ such that: $$q_i \in \arg\max_{q_i} \left[M\cdot\left(\sum_{j=1}^{N} q_j\right)^{-1} - c\right]\cdot q_i $$ which yields the following FOCs: $$\left[M\cdot\left(\sum_{j=1}^{N} q_j\right)^{-1} - c\right] - q_i\cdot M \cdot \left(\sum_{j=1}^{N} q_j\right)^{-2} = 0 \text{ for all } i = 1, 2, ..., N $$
$$\Leftrightarrow \left[M\cdot Q^{-1} - c\right] - q_i\cdot M \cdot Q^{-2} = 0 \text{ for all } i = 1, 2, ..., N $$ If we add all $N$ conditions together, we get: $$ N\cdot\left[M\cdot Q^{-1} - c\right] - Q \cdot M \cdot Q^{-2} = 0 $$ $$\Leftrightarrow N \cdot M\cdot Q^{-1} - N \cdot c - M \cdot Q^{-1} = 0 $$ $$\Leftrightarrow \frac{N - 1}{N} \cdot M\cdot Q^{-1} = c $$
Finally recall that $M\cdot Q^{-1}$ is actually the equilibrium price, meaning that we have: $p = \frac{N}{N-1} c = (1 + \frac{1}{N-1})c $. Consequently, we have: $$Q = \frac{N - 1}{N} \cdot \frac{M}{c}; q_i = \frac{N - 1}{N^2} \cdot \frac{M}{c}; \Pi(N) = \frac{c}{N-1}\cdot \frac{N - 1}{N^2} \cdot \frac{M}{c} = \frac{M}{N^2} $$



\subsubsection{Hotelling/Salop competition}



\section{Static and endogenous models}

\subsection{Sutton's view}



\subsection{Bresnahan and Reiss (1991)}

In a realistic research setting, it is hard to observe strategic variables on a market level (prices, costs, advertising expenses, for all firms). This is the reason why \cite{br_91} innovated on an estimation using mainly superficial market observables: the number of firms (but not the shares) and other market characteristics. The main assumptions of this model rely on a static Sutton model with symmetric firms and free entry.

\subsubsection{Behavioral model}

As in the Sutton-type of models, we solve it backwards, assuming the fixed costs of entry are incurred as sunk in the second period, when firms are choosing production.

Demand in the market $m$ is given by: $$ Q_m = d(Z_m, p)\cdot S(Y_m) $$ where $d(\cdot)$ represents the demand function of a representative consumer and $S(\cdot)$ is the total number of consumers that would buy the product. Note that this demand specification has constant returns to scale: doubling $S$ will double $Q$. Finally, we define the inverse demand curve as $P(Q, Z, Y)$.

Therefore in the second-stage under Cournot competition, each firm solves: $$\max_{q_i}\Pi_{N,m} \equiv P(q_i, q_{-i}, Z_m, Y_m) \cdot q_i - F_N - C(q_i) $$ where $F_N$ is the endogenous sunk cost associated with $N$ firms entering the market. Without loss of generality (we proved a similar result in the Cournot-Sutton context), assume that in equilibrium, quantities will be symmetric such that $q_i = q_j = q^*$ for all $i,j$. Because $N$ is already "chosen" in the second stage, we can write $q^* \equiv d(Z_m, p)\cdot \frac{S(Y_m)}{N}$. Then, using this, the profits of each firm is given by: $$\Pi_{N,m} = P(q^*, q^*, Z_m, Y_m) \cdot q^* - F_N - C(q^*) = \left(P_N - \frac{C(q^*)}{q^*} \right) \cdot d(Z_m, P_N)\cdot \frac{S}{N} - F_N $$ Now, for both the average variable cost $\frac{C(q^*)}{q^*}$ and the fixed cost $F_N$, let's allow them to be additively separable in components, one of them being dependent on the firm only (respectively $AVC$ and $F$) and the other component being endogenous on the number of firms (respectively $b_N$ and $B_N$), such that the total profit of an individual firm, facing a market with $N$ total firms, is defined as: $$\Pi_{N,m} = \left(P_N - AVC(q^*) - b_N \right) \cdot d(Z_m, P_N)\cdot \frac{S}{N} - F - B_N $$

In the first stage, using the free entry assumption, we know that if $N^*$ firms enter a market $m$, it must be that $\Pi_{N^*,m} > 0$ and $\Pi_{N^*+1,m} < 0$. Conversely, we can look at the entry threshold $s_N$, which is the minimum additional demand, not already covered by existing $N-1$ firms that is required in order for the $N$th firm to enter. This threshold is the value of $s_N \equiv S_N/N$ such that: $$\Pi_{N,m} = 0 \Leftrightarrow \frac{S_N}{N} = \frac{F + B_N}{(P_N - AVC_N - b_N)d_N} $$ We can finally also look at the ratio of successive thresholds: $$s_{N+1}/s_N = \frac{F + B_{N+1}}{F + B_N}\frac{(P_N - AVC_{N} - b_N)d_N}{(P_{N+1} - AVC_{N+1} - b_{N+1})d_{N+1}} $$ Thus, in a fully competitive market, a firm would enter each time its cost to enter could be recovered in the market, making $s_{N+1}/s_N$ tend to 1, as $N\to\infty$.

\subsubsection{Empirical strategy}

The threshold equations derived above ask for a lot of observed variables, namely prices, costs (variable and sunk) and more. In reality, it is very difficult to observe all these at the same time for all firms in a market, thus we need a model that could allow for less information, which is the essence of \cite{br_91}.

Consider a reduced-form profit function given by: $$\Pi_{N,m} = \underbrace{\underbrace{S(Y_m, \lambda)}_{\text{size}} \cdot \underbrace{V_{N}(Z_{m}, W_{m}, \alpha, \beta)}_{\text{variable profit}} - \underbrace{F_N(W_{m}, \gamma)}_{\text{fixed}}}_{\bar\Pi_{N,m}} + \epsilon_m $$ which has an endogenous component $\bar\Pi_{N,m}$, and an error-term $\epsilon_m$ that is not dependent on $N$. This should raise some memories to anyone who covered the demand estimation part of the course. In fact, assuming a specification on the error term reveals the model as an ordered probit model, very similar to the logit or multinomial probit models. To see this, consider a market with $N$ firms, this means that: $$ \bar\Pi_{N,m} + \epsilon_m > 0 \text{ and } \bar\Pi_{N+1,m} + \epsilon_m < 0 $$ which is equivalent to: $$ \bar\Pi_{N,m} > \epsilon_m > \bar\Pi_{N+1,m} $$

Assuming $\epsilon_m$ is drawn from an iid normal distribution with mean 0 and variance $\sigma^2$, we have that: $$\prob{N_m} = \begin{cases} \Phi(\bar\Pi_{N,m}) - \Phi(\bar\Pi_{N+1,m}) & \text{ if }N_m > 1 \\
1 - \Phi(\bar\Pi_{2,m}) & \text{ if }N_m = 1
\end{cases} $$

Before going on to the estimation of the model, we need to look at the reduced-form of each component of the profit function:\begin{itemize}
\item The market size includes a combination of market size characteristics $Y_m$, such as the population, the neighbouring population, growth, commuting possibilities, etc.
\item The variable profit per capita is defined as: $V_N = \alpha_1 + X\beta - \sum_{n=2}^{N} \alpha_n $ where:\begin{itemize}
\item $X$ is a vector of relevant economic variables such as demand characteristics for the product ($Z$) and cost-shifters ($W$).
\item $\alpha_n\geq 0$ is an intercept component such that each firm entering a market has a negative effect on profits. 
\end{itemize}
\item The endogenous fixed costs defined as: $F_N = \gamma_1 + \gamma_Lw_L + \sum_{n=2}^{N}\gamma_n\cdot\gamma_n $ 
\end{itemize}

Finally, the estimation is performed using maximum likelihood. In fact, for each observation of the market, we use the probability function defined above. This gives us a likelihood function as a function of all observed variables described in the list above. Taking the log of it yields the log-likelihood to be maximized in order to estimate the parameters of interest.

\subsection{Berry (1992)}

The previous paper was lacking heterogeneity in the sense that its results applies for the case where all firms have the same costs, same continuation payoffs, etc. A step in the direction of allowing some differentiation was made by \cite{berry_92}, where fixed costs can vary across firms. Thus, the behavioral model stays identical, but the profit of firm $k$ can now be divided into a common component $\nu_m(N)$, incurred by all firms in the same way, and an idiosyncratic component $\phi_{m,k}$ applying only to firm $k$. In particular, we have: $$ \Pi_{N,m,k} = \underbrace{X_m\beta - \delta\ln(N) + \rho u_{m,0}}_{\nu_m(N)} + \underbrace{Z_k\alpha + \sqrt{1 - \rho^2}\cdot u_{m,k}}_{\phi_{m,k}} $$ 

Note that in this equation, the combination of the terms $\rho \cdot u_{m,0} + \sqrt{1 - \rho^2}\cdot u_{m,k} $ actually represents the error term, denoted $\epsilon_{m,k}$. In this setting, $\rho$ represents the correlation between error terms across firms in a given market. Finally, \cite{berry_92} assumes that: $$\epsilon_{m,k} \sim N(0, \Sigma) $$ where $\Sigma$ is a matrix with all off-diagonal terms equal to $\rho$.

As in the demand estimation case, this error term is known to the players (the firms) but not to the econometrician. We are still in the static, full information game where all firms know everything about all other firms. However this time we are not in an ordered probit setting since the very structure of the problem is endogenous to the number of firms. To see that, recall the error of the previous model did not include any other subscript than $m$, while this one includes something about $k$ which is intrinsically linked to the number of firms.

Contrary to the previous models, this one can display multiplicity of equilibria. In fact, although the equilibrium number of firms is unique, the exact firms that enter the market can be different. This implies that we lose information on which firms enter a market. For example, the model might predict that in a given market, two firms will ultimately enter, but you could have firms 1 and 2, firms 2 and 3 or firms 1 and 3. In order to control that issue, the author assumes that firms enter in the order of their profitability. Using that assumption, we can simplify the paper and compute the probability of observing $N$ firms in the market as: \begin{align*}
\prob{n_m = N\vert Z_m} & = \prob{\epsilon_{m,1}, ..., \epsilon_{m,K_m} : \sum_{k=1}^{K_m}\mathbb{I}\{\nu_m(N, Z_m) >\phi_{m,k}\} = N } \\
& = \underbrace{\int\hdots\int}_{K_m\text{ times}}\mathbb{I}\left\lbrace\sum_{k=1}^{K_m}\mathbb{I}\{\nu_m(N, Z_m) >\phi_{m,k}\} = N\right\rbrace\D F(\epsilon_{m,1}, ..., \epsilon_{m,K_m}, \theta)
\end{align*} which is a multi-dimensional integral (of $K_m$ dimensions) and as we've seen in the previous two chapters, it is hard to compute. In order to circumvent this computational issue, we could use the same method as in BLP: an inner loop simulating the sample moment of the integral for each value of $\theta$, and an outer loop solving for the value of $\theta$ that minimizes the square distance between the observed and simulated number of firms. This technique is called the Simulated Nonlinear Least Squares (or SNLS).

Formally, the outer loop finds $$\hat\theta = \arg\min_{\theta} \frac{1}{M} \sum_{m} \left( N_m - \E{n_m\vert Z_m, \theta, \epsilon_{m,1}, ..., \epsilon_{m,K_m}} \right)^2 $$ where the expectation term is approximated by its simulated sample average over random iid draws of $(\epsilon_{m,1}, ..., \epsilon_{m,K_m})$. The simulated sample average is defined as: $$\E{n_m\vert Z_m, \theta, \epsilon_{m,1}, ..., \epsilon_{m,K_m}} \approx \frac{1}{S} \sum_{s} n_{m}^{*, s} (\epsilon^s, \theta) $$ $$\text{ where } n_{m}^{*, s} (\epsilon^s, \theta) \equiv \max_{0\leq n \leq K_m} \left\lbrace n : \sum_{k=1}^{K_m}\mathbb{I}\{\nu_m(n, Z_m) > \phi_{m,n}\vert \epsilon^s \}\geq n \right\rbrace $$ or in words, the maximum number of firms $n$, such that the number of firms entering the market with a positive profit is greater than $n$.

To recap the full estimation algorithm, consider these steps:



\section{Two-period models}

\subsection{Stackelberg's model}

Two-period models have been introduced for the first time in entry models, but they have been used in several other settings since, for example in models of investment, contracting, etc. Typically, these models rely on:\begin{itemize}
\item A first period where the agents choose a state variable that will determine the nature of the game in the second period.
\item A second period where the game is played following the state of the world decided on the first.
\end{itemize}
The solution concept used to determine the outcome of the game is called "subgame perfection" which means a Nash equilibrium is chosen at each step played in the game. We also call this equilibrium the Subgame Perfect Nash Equilibrium, or SPNE. In the simple case of a two-period game, the SPNE is quite easy to determine by backwards induction:\begin{itemize}
\item You solve for the NE of the second period game for each potential game played following first period choices.
\item Assuming the outcomes computed above are realized, you solve for the first period choice that is a NE.
\end{itemize}

\subsubsection{Two-period capacity game}

Consider a typical two-period à la Stackelberg game where a firm chooses a level of capital ($K_1 \geq 0$) in the first period, and the second firm chooses its own level ($K_2 \geq 0$) in the second period. A firm that chooses a capital level of $0$ is choosing to stay out of the market. A firm that chooses a level strictly greater than 0 enters the market, although firm 2 would have to pay a fixed cost of entry equal to $F$.

The profits of firm 1 are: $$\pi_1(K_1, K_2) = K_1\cdot (1 - K_1 - K_2) $$
while the profits of firm 2 are: $$ \pi_2(K_1, K_2) = \begin{cases}
K_2\cdot (1 - K_1 - K_2) - F & \text{ if } K_2 > 0 \\
0 & \text{ if } K_2 = 0
\end{cases} $$

In order to solve the game, we proceed by backwards induction. In the second period, we consider the decision of firm 2 over the level $K_2$. It will choose a level $K_2 > 0$ if and only if the profit associated is greater than 0, meaning that: $$ K_2 (1 - K_1 - K_2) - F \geq 0 \Leftrightarrow K_2 (1 - K_1 - K_2) \geq F $$ In that case, firm 2 will choose: $$K_2^* = \arg\max_{K_2} K_2 (1 - K_1 - K_2) - F $$ which is the solution to the following FOC: $$1 - K_1 - 2K_2^* = 0 \Leftrightarrow K_2^* = \frac{1 - K_1}{2} $$ Plugging this solution into the participation condition above, we get that: $$  \frac{1 - K_1}{2} (1 - K_1 - \frac{1 - K_1}{2}) \geq F \Leftrightarrow \frac{(1 - K_1)^2}{4} \geq F \Leftrightarrow K_1 \leq 1 - 2\sqrt{F} $$
and thus the optimal level $K_2^*$ is defined as: $$ K_2^* = \begin{cases}
\frac{1 - K_1}{2} & \text{ if } K_1 \leq 1 - 2\sqrt{F} \\
0 & \text{ if } K_1 > 1 - 2\sqrt{F}
\end{cases} $$

Now that the second period is solved, we can go into the first period. Firm 1 will choose a level $K_1$ to maximize its profits, which are now known to be: $$\pi_1(K_1) = \begin{cases}
K_1 \cdot \left(1 - K_1 - \frac{1 - K_1}{2}\right) & \text{ if } K_1 \leq 1 - 2\sqrt{F} \\
K_1\cdot (1 - K_1) & \text{ if } K_1 > 1 - 2\sqrt{F}
\end{cases} $$ From this perspective, it is clear that, for a given $K_1$, allowing the other player to enter the market is worse than deterring entry. However, deterring entry can only be done for $K_1 > 1 - 2\sqrt{F}$. This creates a discontinuous profit function where deterring entry is always better but can be achieved only at some points that might yield a lower profit than allowing firm 2 to enter. The following graph represents this discrepancy nicely:



\subsection{Empirical work}

 