\section{Lee (2013)}

\begin{itemize}
\item \textbf{Research questions:} What are the effects of vertical integration and exclusive dealing on industry (1) structure and (2) welfare?
\item \textbf{Goals of the paper:} test a model (competitive effects of networks) + measure an effect (effects on structure and welfare).
\item \textbf{Importance of the paper:} New knowledge on network industries (which are more and more common).
\item \textbf{Theoretical foundations:} (1) Two-level demand system (for consoles and software) and (2) developer choice for consoles. \begin{itemize}
\item Strengths: 
\item Shortcomings: No competition between consoles, no console-level strategies.
\end{itemize}
\item \textbf{Empirical strategy:} (1) use dynamic demand and (2) use moment inequalities for developing process.\begin{itemize}
\item Strengths: 
\item Shortcomings: 
\end{itemize}
\item \textbf{Data:} (1) Software sales (genre, release date, exclusivity contract, price, shares, etc.) (2) Hardware sales (shares, prices) and (3) market-level (demographics + survey for heterogeneity).
\begin{itemize}
\item Exogenous: 
\item Endogenous:
\end{itemize}
\item \textbf{Results:} Exclusive arrangements have (1) helped entrants, (2) hurt incumbent and (3) increased differentiation. 
\end{itemize}

\section{Ho (2009)}

\subsection{Discussion}

\begin{itemize}
\item \textbf{Research questions:} (1) How do insurer-provider networks arise? (2) What are the effects of ``star'' hospitals, capacity-constrained hospitals and merged hospitals on the bargaining process?
\item \textbf{Goals of the paper:} test a model (contracting between insurers and providers) + measure an effect (hospital demand or size on bargaining) and maybe slightly interested in policy implications.
\item \textbf{Importance of the paper:} Implications on policy since health providing is so important.
\item \textbf{Theoretical foundations:} (1) Two-level BLP demand system (for insurers and providers) and (2) bargaining game between insurers and providers. \begin{itemize}
\item Strengths: 
\item Shortcomings:
\end{itemize}
\item \textbf{Empirical strategy:} (1) use BLP moments and (2) use moment inequalities for bargaining process (profits from observed network must be higher than reversing decision on one hospital).\begin{itemize}
\item Strengths: 
\item Shortcomings: 
\end{itemize}
\item \textbf{Data:} (1) Hospital-level (size, teaching?, for-profit?, staff size, services, etc.) (2) Plan-level (shares, premia, network size, services included, people included, type of plan, etc.) and (3) market-level (demographics).
\begin{itemize}
\item Exogenous: 
\item Endogenous: contracts, prices, shares.
\end{itemize}
\item \textbf{Results:}  Consumer demand, hospital costs, ``star'' status, capacity constraints and system hospitals are the main features to determine bargaining power of hospitals. 
\end{itemize}