\section{Introduction}

While it was often assumed away in the first semester of IO, we know that firms consider market interactions when choosing entry of products, product positioning, product-line length, etc., thus making these variables endogenous!

Mankiw and Whinston (1986) is among the first paper to consider this issue after Tirole 1980's revolution. The paper shows two main results: (1) entry in a market is endogenous and (2) imperfect competition affects the ``wedge'' between social optimum and private optimum.

\section{Berry and Waldfogel (1999 and 2001)}

\subsection{Discussion}

\begin{itemize}
\item \textbf{Research question:} What are the effects of mergers on product variety?
\item \textbf{Goals of the paper:} test model and measure an effect (concentration, in particular mergers, on product variety).
\item \textbf{Importance of the paper:} Strand of literature on endogeneity of market structure (products), important because regulation can have effects on this.
\item \textbf{Theoretical foundations:}  NOT USED IN EMPIRICAL STRATEGY! Hotelling-style models show ambiguous effects of concentration on product variety. \begin{itemize}
\item Strengths: yields two testable implications.
\item Shortcomings: not used as supply-side in empirical strategy.
\end{itemize}
\item \textbf{Empirical strategy:} use a reduced-form model to correlate changes in concentration with changes in variety using (1) institutional shock and (2) market ``size'' as instruments (number of stations, population).\begin{itemize}
\item Strengths: 
\item Shortcomings: no demand ``adjustment'' to the market due to lacking structural model (tastes?); no welfare study either.
\end{itemize}
\item \textbf{Data:} Market-station-level observations (audience, owner, programming format, size); listening data (AQH)
\begin{itemize}
\item Exogenous: population and listening?
\item Endogenous: changes in concentration, changes in formats (variety)
\end{itemize}
\item \textbf{Results:} Increased concentration reduced station entry without reducing variety (more variety per station).
\end{itemize}

\section{Eizenberg (2014)}



\section{Wollmann (2018)}

